---
title: "multispic diagnostics and results"
output: flexdashboard::flex_dashboard
runtime: shiny
---

```{r setup, echo=FALSE}
library(plotly)
library(viridis)
knitr::opts_chunk$set(echo = FALSE, fig.height = 7, fig.width = 8, cache = FALSE)
```


```{r options}

if (fit$call$K_groups == ~1) {
    K_groups <- NULL
    if (length(unique(fit$landings$species)) == 1) {
        K_label <- "K"
    } else {
        K_label <- "All species"
    }
} else {
    K_groups <- as.formula(fit$call$K_groups)
    K_label <- unique(fit$landings[, all.vars(K_groups)])
}

n_covar <-  length(all.vars(fit$call$pe_covariates))


if (fit$call$species_cor != "none") {
    
    n_spp_cor <- length(fit$par$logit_rho)
    if (n_spp_cor > 1) {
        sp_rho <- sp_nm_mat <- matrix(NA, nrow = nlevels(fit$pop$species), ncol = nlevels(fit$pop$species))
        rownames(sp_rho) <- colnames(sp_rho) <- levels(fit$pop$species)
        sp_rho[lower.tri(sp_rho)] <- inv_logit(fit$par$logit_rho, shift = TRUE)
        sp_rho <- t(sp_rho)
        sp_rho[lower.tri(sp_rho)] <- inv_logit(fit$par$logit_rho, shift = TRUE)
        diag(sp_rho) <- 1
        round(sp_rho, 2)
        for (i in seq(nrow(sp_rho))) {
            for (j in seq(ncol(sp_rho))) {
                sp_nm_mat[i, j] <- paste(levels(fit$pop$species)[i], "-", levels(fit$pop$species)[j])
            }
        }
        species_cor_label <- sp_nm_mat[lower.tri(sp_nm_mat)]
    } else {
        species_cor_label <- "All species"
    }
    
} else {
    
    n_spp_cor <- 0
    species_cor_label <- "N/A"
    
}

n_time_cor <- as.numeric(fit$call$temporal_cor == "ar1")
temporal_cor_label <- ifelse(n_time_cor == 1, "All species", "N/A")

add_labels <- function(p, xlab = "", ylab = "", log_linear_button = TRUE) {
    
    if (log_linear_button) {
        ## Modified from https://stackoverflow.com/a/60551263/4055438
        buttons <- list(list(
            active = 0,
            type = "buttons",
            buttons= list(
                list(label = 'linear',
                     method = 'relayout',
                     args = list(list(yaxis = list(type = 'linear', title = ylab)))),
                list(label = 'log',
                     method = 'relayout', 
                     args = list(list(yaxis = list(type = 'log', title = ylab)))))))
    } else {
        buttons <- NULL
    }
    
    layout(p, 
           xaxis = list(title = xlab),
           yaxis = list(title = ylab),
           updatemenus = buttons)
    
}


```


Inputs
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Index

```{r}
fit$index %>%
    plot_ly() %>%
    add_trace(x = ~year, y = ~index, color = ~survey,
              colors = viridis::viridis(100), mode = "markers+lines",
              type = "scatter") %>% 
    add_labels(xlab = "Year", ylab = "Index")
```


### Landings

```{r}
fit$landings %>%
    plot_ly() %>%
    add_trace(x = ~year, y = ~landings, color = ~species,
              colors = viridis::viridis(100), mode = "markers+lines",
              type = "scatter") %>% 
    add_labels(xlab = "Year", ylab = "Landings")
```


### Covariates

`r if (n_covar == 0) "N/A; no process error covariates supplied"`
`r if (n_covar > 1) "Plotting of more than one covariate not yet implemented"`

```{r, eval = n_covar == 1}
covariates <- unique(fit$landings[, c("year", all.vars(fit$call$pe_covariates)), drop = FALSE])
names(covariates) <- c("year", "covariate")
covariates %>% 
    plot_ly(x = ~year, y = ~covariate) %>% 
    add_trace(mode = "markers+lines") %>% 
    layout(xaxis = list(title = "Year"),
           yaxis = list(title = all.vars(fit$call$pe_covariates)))
```



Residuals
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Residuals ~ predicted value

```{r}
p <- fit$index %>%
    plot_ly(color = ~species, colors = viridis::viridis(100)) %>% 
    layout(yaxis = list(title = "Standardized residuals"))
p %>% add_markers(x = ~log(pred), y = ~std_res)
```

### ~ year

```{r}
p %>% add_markers(x = ~year, y = ~std_res)
```

### ~ survey

```{r}
p %>% add_markers(x = ~survey, y = ~std_res)
```

### ~ species

```{r}
p %>% add_markers(x = ~species, y = ~std_res)
```


Parameters
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Correlation

```{r}

sd_rep <- fit$sd_rep
dsd <- sqrt(diag(sd_rep$cov.fixed))
cor_mat <- diag(1 / dsd) %*% sd_rep$cov.fixed %*% diag(1 / dsd)
rownames(cor_mat) <- paste0(seq(nrow(cor_mat)), ":", names(dsd))
colnames(cor_mat) <- paste0(seq(nrow(cor_mat)), ":", names(dsd))
# corrplot::corrplot.mixed(cor_mat, diag = "n", lower = "ellipse", upper = "number")
plot_ly(x = rownames(cor_mat), y = colnames(cor_mat), z = cor_mat) %>% 
    add_heatmap(colors = c("#B2182B", "white", "#2166AC")) %>% 
    colorbar(limits = c(-1.001, 1.001))

```

### log(K)

```{r}
plot_post(prior_mean = fit$par$mean_log_K,
          prior_sd = exp(fit$par$log_sd_log_K),
          post_mean = fit$par$log_K,
          post_sd = fit$se$log_K,
          post_names = K_label,
          xlab = "log(K)",
          show_prior = fit$call$log_K_option$option == "normal_prior")
```

### log(r)

```{r}
plot_post(prior_mean = fit$par$mean_log_r,
          prior_sd = exp(fit$par$log_sd_log_r),
          post_mean = fit$par$log_r,
          post_sd = fit$se$log_r,
          post_names = levels(fit$landings$species),
          xlab = "log(r)",
          show_prior = fit$call$log_r_option$option == "normal_prior")
```


### log(B0)

```{r}
plot_post(prior_mean = fit$par$mean_log_B0,
          prior_sd = exp(fit$par$log_sd_log_B0),
          post_mean = fit$par$log_B0,
          post_sd = fit$se$log_B0,
          post_names = levels(fit$landings$species),
          xlab = "log(B0)",
          show_prior = fit$call$log_B0_option$option == "normal_prior")
```

### log(SD<sub>B</sub>)

```{r}
plot_post(prior_mean = fit$par$mean_log_sd_B,
          prior_sd = exp(fit$par$log_sd_log_sd_B),
          post_mean = fit$par$log_sd_B,
          post_sd = fit$se$log_sd_B,
          post_names = levels(fit$landings$species),
          xlab = "log(SD<sub>B</sub>)",
          show_prior = fit$call$log_sd_B_option$option == "normal_prior")

```


### logit(rho)

`r if (n_spp_cor == 0) "N/A; correlation of process errors across species was not estimated."`

```{r, eval = n_spp_cor > 0}
plot_post(prior_mean = fit$par$mean_logit_rho,
          prior_sd = exp(fit$par$log_sd_logit_rho),
          post_mean = fit$par$logit_rho,
          post_sd = fit$se$logit_rho,
          post_names = species_cor_label,
          xlab = "logit(rho)", # trans_fun = function(x) inv_logit(x, shift = TRUE),
          show_prior = fit$call$logit_rho_option$option == "normal_prior")
```



### logit(phi)

`r if (n_time_cor == 0) "N/A; correlation of process errors across time was not estimated."`

```{r, eval = n_time_cor == 1}
plot_post(prior_mean = fit$obj$env$data$mean_logit_phi,
          prior_sd = fit$obj$env$data$sd_logit_phi,
          post_mean = fit$par$logit_phi,
          post_sd = fit$se$logit_phi,
          post_names = "logit(phi)",
          xlab = "logit(phi)", # trans_fun = inv_logit,
          show_prior = fit$call$logit_phi_option$option == "normal_prior")

```


### log(q)

```{r}
plot_post(prior_mean = fit$par$mean_log_q,
          prior_sd = exp(fit$par$log_sd_log_q),
          post_mean = fit$par$log_q,
          post_sd = fit$se$log_q,
          post_names = levels(fit$index$survey),
          xlab = "log(q)",
          show_prior = fit$call$log_q_option$option == "normal_prior")
```


### log(SD<sub>I</sub>)

```{r}
plot_post(prior_mean = fit$par$mean_log_sd_I,
          prior_sd = exp(fit$par$log_sd_log_sd_I),
          post_mean = fit$par$log_sd_I,
          post_sd = fit$se$log_sd_I,
          post_names = levels(fit$index$survey),
          xlab = "log(SD<sub>I</sub>)",
          show_prior = fit$call$log_sd_I_option$option == "normal_prior")

```

### Estimates (plot)

```{r}

name_par <- c("log_K", "log_r", "log_B0", "log_sd_B",
              "logit_rho", "logit_phi", "log_q", "log_sd_I",
              "pe_betas")
par_fun <- function(nm, select = "par") {
    if (grepl("beta", nm)) return(fit[[select]][[nm]])
    if (grepl("log_", nm)) return(exp(fit[[select]][[nm]]))
    if (nm == "logit_rho") return(inv_logit(fit[[select]][[nm]], shift = TRUE))
    if (nm == "logit_phi") return(inv_logit(fit[[select]][[nm]], shift = FALSE))
}
named_par <- lapply(name_par, par_fun, select = "par")
named_par_cv <- lapply(name_par, par_fun, select = "se")
named_par_lwr <- lapply(name_par, par_fun, select = "par_lwr")
named_par_upr <- lapply(name_par, par_fun, select = "par_upr")
names(named_par) <- gsub("log_|logit_", "", name_par)
names(named_par$q) <- levels(fit$index$survey)
names(named_par$sd_I) <- levels(fit$index$survey)
names(named_par$K) <- K_label
names(named_par$r) <- levels(fit$index$species)
names(named_par$sd_B) <- levels(fit$index$species)
names(named_par$B0) <- levels(fit$index$species)
names(named_par$phi) <- temporal_cor_label
names(named_par$rho) <- species_cor_label

par_tab <- data.frame(parameter = rep(names(named_par), sapply(named_par, length)),
                      group = names(unlist(unname(named_par))),
                      estimate = unlist(unname(named_par)),
                      CV = unlist(named_par_cv),
                      lower = unlist(named_par_lwr),
                      upper = unlist(named_par_upr))
par_tab <- na.omit(par_tab) # will drop par lacking CV estimates (i.e. not estimated)

par_tab %>% 
    plot_ly(color = ~parameter, colors = viridis::viridis(100),
            legendgroup = ~parameter, y = ~paste(parameter, "-", group)) %>% 
    add_segments(x = ~lower, xend = ~upper, yend = ~paste(parameter, "-", group), 
                 showlegend = FALSE) %>% 
    add_markers(x = ~estimate) %>% 
    layout(xaxis = list(title = ""), yaxis = list(title = ""))

```


### Estimates (table)

```{r}
par_tab$parameter[duplicated(par_tab$parameter)] <- ""
knitr::kable(par_tab, digits = 3, row.names = FALSE)
```



Population trends
================================================================================

```{r}

```


