---
title: "multispic diagnostics and results"
output:
  flexdashboard::flex_dashboard:
    theme: 
      version: 4
      bootswatch: litera
runtime: shiny
---

```{r setup, echo=FALSE}
library(plotly)
library(viridis)
knitr::opts_chunk$set(echo = FALSE, fig.height = 7, fig.width = 8, cache = FALSE)
```


```{r options}

if (fit$call$K_groups == ~1) {
  K_groups <- NULL
  if (length(unique(fit$landings$species)) == 1) {
    K_label <- "K"
  } else {
    K_label <- "All species"
  }
} else {
  K_groups <- as.formula(fit$call$K_groups)
  K_label <- unique(fit$landings[, all.vars(K_groups)])
}

n_pe_covar <-  length(all.vars(fit$call$pe_covariates))
n_K_covar <-  length(all.vars(fit$call$K_covariates))

if (fit$call$species_cor != "none") {
  
  n_spp_cor <- length(fit$par$logit_rho)
  if (n_spp_cor > 1) {
    sp_rho <- sp_nm_mat <- matrix(NA, nrow = nlevels(fit$pop$species), ncol = nlevels(fit$pop$species))
    rownames(sp_rho) <- colnames(sp_rho) <- levels(fit$pop$species)
    sp_rho[lower.tri(sp_rho)] <- inv_logit(fit$par$logit_rho, shift = TRUE)
    sp_rho <- t(sp_rho)
    sp_rho[lower.tri(sp_rho)] <- inv_logit(fit$par$logit_rho, shift = TRUE)
    diag(sp_rho) <- 1
    round(sp_rho, 2)
    for (i in seq(nrow(sp_rho))) {
      for (j in seq(ncol(sp_rho))) {
        sp_nm_mat[i, j] <- paste(levels(fit$pop$species)[i], "-", levels(fit$pop$species)[j])
      }
    }
    species_cor_label <- sp_nm_mat[lower.tri(sp_nm_mat)]
  } else {
    species_cor_label <- "All species"
  }
  
} else {
  
  n_spp_cor <- 0
  species_cor_label <- "N/A"
  
}

n_time_cor <- as.numeric(fit$call$temporal_cor == "ar1")
temporal_cor_label <- ifelse(n_time_cor == 1, "All species", "N/A")

add_labels <- function(p, xlab = "", ylab = "", log_linear_button = TRUE) {
  
  if (log_linear_button) {
    ## Modified from https://stackoverflow.com/a/60551263/4055438
    buttons <- list(list(
      active = 0,
      type = "buttons",
      buttons= list(
        list(label = 'linear',
             method = 'relayout',
             args = list(list(yaxis = list(type = 'linear', title = ylab)))),
        list(label = 'log',
             method = 'relayout', 
             args = list(list(yaxis = list(type = 'log', title = ylab)))))))
  } else {
    buttons <- NULL
  }
  
  layout(p, 
         xaxis = list(title = xlab),
         yaxis = list(title = ylab),
         updatemenus = buttons)
  
}


```


Inputs
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Index

```{r}
fit$index %>%
  plot_ly() %>%
  add_trace(x = ~year, y = ~index, color = ~survey,
            colors = viridis::viridis(100), mode = "markers+lines",
            type = "scatter") %>% 
  add_labels(xlab = "Year", ylab = "Index")
```


### Landings

```{r}
fit$landings %>%
  plot_ly() %>%
  add_trace(x = ~year, y = ~landings, color = ~species,
            colors = viridis::viridis(100), mode = "markers+lines",
            type = "scatter") %>% 
  add_labels(xlab = "Year", ylab = "Landings")
```


### Covariates

`r if (n_pe_covar == 0 && n_K_covar == 0) "N/A; no process error or K covariates supplied"`
`r if ((n_pe_covar + n_K_covar) > 1) "Plotting of more than one covariate not yet implemented"`

```{r, eval = (n_pe_covar + n_K_covar) == 1}
if (n_pe_covar == 1) {
  covar_name <- all.vars(fit$call$pe_covariates)
  covariates <- unique(fit$landings[, c("year", covar_name), drop = FALSE])
}
if (n_K_covar == 1) {
  covar_name <- all.vars(fit$call$K_covariates)
  covariates <- unique(fit$landings[, c("year", covar_name), drop = FALSE])
}
names(covariates) <- c("year", "covariate")
covariates %>% 
  plot_ly(x = ~year, y = ~covariate) %>% 
  add_lines() %>% 
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = covar_name))
```


Residuals
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Residuals ~ predicted value

```{r}
p <- fit$index %>%
  plot_ly(color = ~species, colors = viridis::viridis(100)) %>% 
  layout(yaxis = list(title = "Standardized residuals"))
p %>% add_markers(x = ~log(pred), y = ~std_res)
```

### ~ year

```{r}
p %>% add_markers(x = ~year, y = ~std_res)
```

### ~ survey

```{r}
p %>% add_markers(x = ~survey, y = ~std_res)
```

### ~ species

```{r}
p %>% add_markers(x = ~species, y = ~std_res)
```


Parameters
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Correlation

```{r}

sd_rep <- fit$sd_rep
dsd <- sqrt(diag(sd_rep$cov.fixed))
cor_mat <- diag(1 / dsd) %*% sd_rep$cov.fixed %*% diag(1 / dsd)
rownames(cor_mat) <- paste0(seq(nrow(cor_mat)), ":", names(dsd))
colnames(cor_mat) <- paste0(seq(nrow(cor_mat)), ":", names(dsd))
# corrplot::corrplot.mixed(cor_mat, diag = "n", lower = "ellipse", upper = "number")
plot_ly(x = rownames(cor_mat), y = colnames(cor_mat), z = cor_mat) %>% 
  add_heatmap(colors = c("#B2182B", "white", "#2166AC")) %>% 
  colorbar(limits = c(-1.001, 1.001)) %>% 
  layout(yaxis = list(scaleanchor = "x", scaleratio = 1),
         xaxis = list(constrain = "domain"))

```

### log(K)

```{r}
plot_post(prior_mean = fit$par$mean_log_K,
          prior_sd = exp(fit$par$log_sd_log_K),
          post_mean = fit$par$log_K,
          post_sd = fit$se$log_K,
          post_names = K_label,
          xlab = "log(K)",
          show_prior = fit$call$log_K_option$option == "normal_prior")
```

### log(r)

```{r}
plot_post(prior_mean = fit$par$mean_log_r,
          prior_sd = exp(fit$par$log_sd_log_r),
          post_mean = fit$par$log_r,
          post_sd = fit$se$log_r,
          post_names = levels(fit$landings$species),
          xlab = "log(r)",
          show_prior = fit$call$log_r_option$option == "normal_prior")
```


### log(B0)

```{r}
plot_post(prior_mean = fit$par$mean_log_B0,
          prior_sd = exp(fit$par$log_sd_log_B0),
          post_mean = fit$par$log_B0,
          post_sd = fit$se$log_B0,
          post_names = levels(fit$landings$species),
          xlab = "log(B0)",
          show_prior = fit$call$log_B0_option$option == "normal_prior")
```

### log(SD<sub>B</sub>)

```{r}
plot_post(prior_mean = fit$par$mean_log_sd_B,
          prior_sd = exp(fit$par$log_sd_log_sd_B),
          post_mean = fit$par$log_sd_B,
          post_sd = fit$se$log_sd_B,
          post_names = levels(fit$landings$species),
          xlab = "log(SD<sub>B</sub>)",
          show_prior = fit$call$log_sd_B_option$option == "normal_prior")

```


### logit(rho)

`r if (n_spp_cor == 0) "N/A; correlation of process errors across species was not estimated."`

```{r, eval = n_spp_cor > 0}
plot_post(prior_mean = fit$par$mean_logit_rho,
          prior_sd = exp(fit$par$log_sd_logit_rho),
          post_mean = fit$par$logit_rho,
          post_sd = fit$se$logit_rho,
          post_names = species_cor_label,
          xlab = "logit(rho)", # trans_fun = function(x) inv_logit(x, shift = TRUE),
          show_prior = fit$call$logit_rho_option$option == "normal_prior") 
```



### logit(phi)

`r if (n_time_cor == 0) "N/A; correlation of process errors across time was not estimated."`

```{r, eval = n_time_cor == 1}
plot_post(prior_mean = fit$tmb_dat$mean_logit_phi,
          prior_sd = fit$tmb_dat$sd_logit_phi,
          post_mean = fit$par$logit_phi,
          post_sd = fit$se$logit_phi,
          post_names = "logit(phi)",
          xlab = "logit(phi)", # trans_fun = inv_logit,
          show_prior = fit$call$logit_phi_option$option == "normal_prior")

```


### log(q)

```{r}
plot_post(prior_mean = fit$par$mean_log_q,
          prior_sd = exp(fit$par$log_sd_log_q),
          post_mean = fit$par$log_q,
          post_sd = fit$se$log_q,
          post_names = levels(fit$index$survey),
          xlab = "log(q)",
          show_prior = fit$call$log_q_option$option == "normal_prior")
```


### log(SD<sub>I</sub>)

```{r}
plot_post(prior_mean = fit$par$mean_log_sd_I,
          prior_sd = exp(fit$par$log_sd_log_sd_I),
          post_mean = fit$par$log_sd_I,
          post_sd = fit$se$log_sd_I,
          post_names = levels(fit$index$survey),
          xlab = "log(SD<sub>I</sub>)",
          show_prior = fit$call$log_sd_I_option$option == "normal_prior")

```

### beta<sub>pe</sub>

`r if (n_pe_covar == 0) "N/A; no process error covariates supplied"`

```{r, eval = n_pe_covar >= 1}
plot_post(prior_mean = fit$tmb_dat$mean_pe_betas,
          prior_sd = fit$tmb_dat$sd_pe_betas,
          post_mean = fit$par$pe_betas,
          post_sd = fit$se$pe_betas,
          post_names = "pe_betas",
          xlab = "pe_betas",
          show_prior = fit$call$pe_betas_option$option == "normal_prior")
```

### beta<sub>K</sub>

`r if (n_K_covar == 0) "N/A; no K covariates supplied"`

```{r, eval = n_K_covar >= 1}
plot_post(prior_mean = fit$tmb_dat$mean_K_betas,
          prior_sd = fit$tmb_dat$sd_K_betas,
          post_mean = fit$par$K_betas,
          post_sd = fit$se$K_betas,
          post_names = "K_betas",
          xlab = "K_betas",
          show_prior = fit$call$K_betas_option$option == "normal_prior")
```


### Estimates (plot)

```{r}

name_par <- c("log_K", "log_r", "log_B0", "log_sd_B",
              "logit_rho", "logit_phi", "log_q", "log_sd_I",
              "pe_betas", "K_betas")
par_fun <- function(nm, select = "par") {
  if (grepl("beta", nm)) return(fit[[select]][[nm]])
  if (grepl("log_", nm)) return(exp(fit[[select]][[nm]]))
  if (nm == "logit_rho") return(inv_logit(fit[[select]][[nm]], shift = TRUE))
  if (nm == "logit_phi") return(inv_logit(fit[[select]][[nm]], shift = FALSE))
}
named_par <- lapply(name_par, par_fun, select = "par")
named_par_cv <- lapply(name_par, par_fun, select = "se")
named_par_lwr <- lapply(name_par, par_fun, select = "par_lwr")
named_par_upr <- lapply(name_par, par_fun, select = "par_upr")
names(named_par) <- gsub("log_|logit_", "", name_par)
names(named_par$q) <- levels(fit$index$survey)
names(named_par$sd_I) <- levels(fit$index$survey)
names(named_par$K) <- K_label
names(named_par$r) <- levels(fit$index$species)
names(named_par$sd_B) <- levels(fit$index$species)
names(named_par$B0) <- levels(fit$index$species)
names(named_par$phi) <- temporal_cor_label
names(named_par$rho) <- species_cor_label
names(named_par$pe_betas) <- colnames(fit$tmb_dat$pe_covariates)
names(named_par$K_betas) <- colnames(fit$tmb_dat$K_covariates)

par_tab <- data.frame(parameter = rep(names(named_par), sapply(named_par, length)),
                      group = names(unlist(unname(named_par))),
                      estimate = unlist(unname(named_par)),
                      CV = unlist(named_par_cv),
                      lower = unlist(named_par_lwr),
                      upper = unlist(named_par_upr))
par_tab <- na.omit(par_tab) # will drop par lacking CV estimates (i.e. not estimated)

par_tab %>% 
  plot_ly(color = ~parameter, colors = viridis::viridis(100),
          legendgroup = ~parameter, y = ~paste(parameter, "-", group)) %>% 
  add_segments(x = ~lower, xend = ~upper, yend = ~paste(parameter, "-", group), 
               showlegend = FALSE) %>% 
  add_markers(x = ~estimate) %>% 
  layout(xaxis = list(title = ""), yaxis = list(title = ""))

```


### Estimates (table)

```{r}
par_tab$parameter[duplicated(par_tab$parameter)] <- ""
knitr::kable(par_tab, digits = 3, row.names = FALSE)
```



Population trends
================================================================================


Row {.tabset}
--------------------------------------------------------------------------------

### Observed and predicted index

```{r}
fit$index %>%
  plot_ly(x = ~year, color = ~survey, colors = viridis::viridis(100),
          legendgroup = ~survey) %>%
  add_ribbons(ymin = ~pred_lwr, ymax = ~pred_upr, line = list(width = 0),
              alpha = 0.2, showlegend = FALSE) %>%
  add_lines(y = ~pred) %>%
  add_markers(y = ~index, showlegend = FALSE) %>% 
  add_labels(xlab = "Year", ylab = "Index")
```


### Process error

```{r}
p <- fit$pop %>%
  plot_ly(color = ~species, colors = viridis::viridis(100))
p %>% add_lines(x = ~year, y = ~pe) %>% 
  add_labels(xlab = "Year", ylab = "Process error")
```

`r if (n_pe_covar >= 1) "> Includes covariate effects"`


### Standardized process error

```{r}
p <- fit$pop %>%
  plot_ly(color = ~species, colors = viridis::viridis(100))
p %>% add_lines(x = ~year, y = ~log_std_res_pe) %>% 
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = "Standardized process error"))
```

`r if (n_pe_covar >= 1) "> Standardized process error = residual process error, in log space, not explained by the covariate effects divided by the standard deviation of the process." else "> Standardized process error = process error, in log space, divided by the standard deviation of the process."`


### Process error correlation

`r if (n_spp_cor == 0) "N/A; correlation of process errors across species was not estimated."`
`r if (n_spp_cor == 1) "N/A; matrix of correlation was not estimated as rho was coupled across species."`

```{r, eval = n_spp_cor > 1}
plot_ly(x = rownames(sp_rho), y = rownames(sp_rho), z = ~sp_rho,
        colors = c("#B2182B", "white", "#2166AC")) %>%
  add_heatmap() %>% colorbar(limits = c(-1.001, 1.001), title = "Ï") %>% 
  layout(yaxis = list(scaleanchor = "x", scaleratio = 1),
         xaxis = list(constrain = "domain"))
```

`r if (n_spp_cor > 1 && n_pe_covar >= 1) "> Represents correlation in residual process error not explained by the covariate effects."`


### Biomass

```{r}

tots <- fit$tot_pop
if (is.null(K_groups)) {
  tots$group <- "K"
} else {
  tots$group <- paste0("K", "-", tots[, all.vars(K_groups)])
}

fit$pop %>%
  ungroup() %>% 
  plot_ly(x = ~year, color = ~species, colors = viridis::viridis(100),
          legendgroup = ~species) %>%
  add_ribbons(ymin = ~B_lwr, ymax = ~B_upr, line = list(width = 0),
              alpha = 0.2, showlegend = FALSE) %>%
  add_lines(y = ~B) %>% 
  add_lines(data = tots, x = ~year, y = ~K, 
            linetype = ~group, color = I("black"), inherit = FALSE) %>% 
  add_labels(xlab = "Year", ylab = "Biomass")


```

### Total biomass

```{r}
three_cols <- viridis::viridis(3)
fit$tot_pop %>%
  plot_ly(x = ~year, frame = K_groups) %>%
  add_ribbons(ymin = ~B_lwr, ymax = ~B_upr, line = list(width = 0),
              alpha = 0.2, showlegend = FALSE, legendgroup = "B",
              color = I(three_cols[2])) %>%
  add_lines(y = ~B, name = "B", color = I(three_cols[2]),
            legendgroup = "B") %>%
  add_lines(y = ~K, name = "K", legendgroup = "K",
            linetype = I(1), color = I(three_cols[1])) %>%
  add_lines(y = ~K_lwr, legendgroup = "K", showlegend = FALSE,
            linetype = I(3), color = I(three_cols[1]), size = I(1)) %>%
  add_lines(y = ~K_upr, legendgroup = "K", showlegend = FALSE,
            linetype = I(3), color = I(three_cols[1]), size = I(1)) %>%
  add_lines(y = ~landings, name = "L", color = I(three_cols[3])) %>%
  add_labels(xlab = "Year", ylab = "Biomass")
```


### Harvest rate

```{r}
fit$pop %>%
  plot_ly(x = ~year, color = ~species, colors = viridis::viridis(100),
          legendgroup = ~species) %>%
  add_ribbons(ymin = ~F_lwr, ymax = ~F_upr, line = list(width = 0),
              alpha = 0.2, showlegend = FALSE) %>%
  add_lines(y = ~F) %>%
  add_labels(xlab = "Year", ylab = "Harvest Rate")
```


